Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics

### 2022-02-23 Thesis/_book/thesis.tex
add regressor distribution

Based on the answer to the above questions, four major approaches to pursue AI have been established. These approaches can be summarized into a two by two table as shown in figure \ref{fig:twodimai}, where the row is ``Human'' vs.~``Rational'', and the column is ``Behaviour'' vs.~``Thought''. Positioning at the top right cell, the \textbf{rational agent approach} aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved. In contrast, the \textbf{``laws of thought'' approach} focus on understanding the logic behind the rationality. Closely related to \textbf{cognitive science}, the \textbf{cognitive modelling approach} attempts to express theories of human cognition as computer program to mimic the thought process of human. Lastly, the \textbf{Turing test approach} is built upon the famous \textbf{Turing test} proposed by \textcite{turing_computing_1950}. The test can be roughly described as, whether a human can distinguish another human from a computer with written communications only. To pass the test, several capabilities of computer are required. This includes \textbf{natural language processing} for communication with human, \textbf{knowledge representation} for encoding knowledge, \textbf{automated reasoning} for derivation of conclusions and \textbf{machine learning} for improving AI automatically through experience and data. Some researchers argued that written communication is insufficient to demonstrate intelligence, and some degree of physical simulation of a person is still necessary. One such example is the \textbf{total Turing test} proposed by \textcite{harnad_other_1991}. It adds \(3\) new requirements to the list, including \textbf{computer vision}, \textbf{speech recognition} and \textbf{robotics}, which are response for interactions with the physical world. Notably, all \(7\) required capabilities have become major subfields of AI today. And their development has made AI one of the fastest-growing fields in the 21st century \autocite{russell_artificial_2002}.

@@

In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries \autocite{buja_statistical_2009}. For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis. The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in figure \ref{fig:nonlinearityheter}.

@@

From the figure \ref{fig:cubic1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes. Figure \ref{fig:cubic2} shows that with fixed \(a\), \(b\) and \(\sigma\), \(c\) is controlling the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\). If we move \(c\) toward \(0\) or \(2\), one variable will donminate another, which will mitigate some joint effects and resemble a typical cubic function. In figure \ref{fig:cubic3}, \(a\), \(b\) and \(c\) is held constant, and \(\sigma\) is controlling the noises around the expected values. As \(\sigma\) decreases, the underlying shape shows up.

@@

The cubic model involves the use of both \(\boldsymbol{X}\) and \(\boldsymbol{Z}\). In the simulation, \(X_i\), \(i=1,...,n\), had equal chance to follow one of the following distributions: \(N(0, 0.09)\), \(U(-1, 1)\), \(Lognormal(0,0.36)\) and \(-Lognormal(0,0.36)\).

Normal and uniform distributions are

>>>>

Based on the answer to the above questions, four major approaches to pursue AI have been established. These approaches can be summarized into a two by two table as shown in Figure \ref{fig:twodimai}, where the row is ``Human'' vs.~``Rational'', and the column is ``Behaviour'' vs.~``Thought''. Positioning at the top right cell, the \textbf{rational agent approach} aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved. In contrast, the \textbf{``laws of thought'' approach} focus on understanding the logic behind the rationality. Closely related to \textbf{cognitive science}, the \textbf{cognitive modelling approach} attempts to express theories of human cognition as computer program to mimic the thought process of human. Lastly, the \textbf{Turing test approach} is built upon the famous \textbf{Turing test} proposed by \textcite{turing_computing_1950}. The test can be roughly described as, whether a human can distinguish another human from a computer with written communications only. To pass the test, several capabilities of computer are required. This includes \textbf{natural language processing} for communication with human, \textbf{knowledge representation} for encoding knowledge, \textbf{automated reasoning} for derivation of conclusions and \textbf{machine learning} for improving AI automatically through experience and data. Some researchers argued that written communication is insufficient to demonstrate intelligence, and some degree of physical simulation of a person is still necessary. One such example is the \textbf{total Turing test} proposed by \textcite{harnad_other_1991}. It adds \(3\) new requirements to the list, including \textbf{computer vision}, \textbf{speech recognition} and \textbf{robotics}, which are response for interactions with the physical world. Notably, all \(7\) required capabilities have become major subfields of AI today. And their development has made AI one of the fastest-growing fields in the 21st century \autocite{russell_artificial_2002}.

@@

In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries \autocite{buja_statistical_2009}. For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis. The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \ref{fig:nonlinearityheter}.

@@

From the Figure \ref{fig:cubic1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes. Figure \ref{fig:cubic2} shows that with fixed \(a\), \(b\) and \(\sigma\), \(c\) is controlling the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\). If we move \(c\) toward \(0\) or \(2\), one variable will donminate another, which will mitigate some joint effects and resemble a typical cubic function. In Figure \ref{fig:cubic3}, \(a\), \(b\) and \(c\) is held constant, and \(\sigma\) is controlling the noises around the expected values. As \(\sigma\) decreases, the underlying shape shows up.

@@

The cubic model involved the use of both \(\boldsymbol{X}\) and \(\boldsymbol{Z}\). In the simulation, \(X_i\), \(i=1,...,n\), had equal chance to follow one of the following distributions: \(N(0, 0.09)\), \(U(-1, 1)\), \(Lognormal(0,0.36)/3 - 1\) and \(-Lognormal(0,0.36)/3 + 1\). Uniform and normal distribution were symmetric and common. Adjusted lognormal distribution and adjusted negative lognormal distribution provided right-skewed and left-skewed density respectively. These distributions were chosen such that most the realizations will fall between \(-1\) and \(1\).

The distribution of \(Z_i\), \(i=1,...,n\), had \(50\)\% chance to be a uniform distribution ranged from \(-1\) to \(1\), and \(50\)\% chance to be a discrete uniform distribution with \(z_n\) outcomes simulated from a uniform distribution ranged from \(-1\) to \(1\). \(z_n\) itself was a discrete uniform distribution with outcomes \(\{3,4,5,6,7,8,9,10\}\), which defined the number of possible values \(Z_i\) could take. As shown in Figure \ref{fig:discreteex}, this setup would create discreteness in residual plot, which could enrich the pool of visual patterns.

\begin{figure}
\centering
\includegraphics{thesis_files/figure-latex/discreteex-1.pdf}
\caption{\label{fig:discreteex}Discreteness in residuals created by using a discrete uniform random variable as one of the regressors in a cubic model. For each residual plot of the lineup, there are three clusters because the number of possible values the regressor can take is three.}
\end{figure}

\(\boldsymbol{X}\) used in the heteroskedasticity model was a combination of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) used in the cubic model. It could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted lognormal distribution, adjusted negative lognormal distribution and discrete uniform distribution.

\hypertarget{experiment-i}{%
\section{Experiment I}\label{experiment-i}}

\hypertarget{experiment-ii}{%
\section{Experiment II}\label{experiment-ii}}
